---
output:
  pdf_document: default
  html_document: default
---
# Introduction to Data Science HW 4
##### Copyright Jeffrey Stanton, Jeffrey Saltz, and Jasmina Tacheva


```{r}
# Enter your name here: Cy Seeley
```

### Attribution statement: (choose only one and delete the rest)


```{r}
# 1. I did this homework by myself, with help from the book and the professor.

```

Reminders of things to practice from previous weeks: <br>
Descriptive statistics: 	mean( ) max( ) min( )<br>
Coerce to numeric:		as.numeric( )<br>

## Part 1: Use the Starter Code 

Below, I have provided a starter file to help you. <br>

Each of these lines of code **must be commented** (the comment must that explains what is going on, so that I know you understand the code and results).


```{r}
#This code is simply calling the jsonlite package in R
library(jsonlite)
#This is creating an object called dataset which uses the url function and follows a given url and brings back a dataset we will use later on
dataset <- url("https://intro-datascience.s3.us-east-2.amazonaws.com/role.json")
#This line of code is converting our URL dataset to a dataframe which is usable in R
readlines <- jsonlite::fromJSON(dataset)
#This code just trims the original data set to something a bit smaller and easier to work with
df <- readlines$objects$person
```

A.	Explore the **df** dataframe (e.g., using head() or whatever you think is best).

head(df)

```{r}


```

B.	Explain the dataset <br>
o	What is the dataset about? <br>
#This is a dataset about different congressman and contains variables such as their name and social media id's
o	How many rows are there and what does a row represent?<br>
#There is 100 rows and each of them represents a different congressman
o	How many columns and what does each column represent?<br>
#There is 17 different columns and each one is a different varaible such as their name, social media id's, and genders



```{r}

```

C. What does running this line of code do? Explain in a comment:
#it creates a list of the years in which each congressman was born

```{r}
vals <- substr(df$birthday,1,4)
```

D. Create a new attribute 'age' - how old the person is 
**Hint:** You may need to convert it to numeric first.

valsnum <- as.numeric(vals)
thisyear <- format(Sys.Date(), "%Y")
thisyearnum <- as.numeric(thisyear)
age <- thisyearnum - valsnum

```{r}

```

E. Create a function that reads in the role json dataset, and adds the age attribute to the dataframe, and returns that dataframe


```


```{r}
addnew <- function() {
  library(jsonlite)
  dataset <- url("https://intro-datascience.s3.us-east-2.amazonaws.com/role.json")
  readlines <- jsonlite::fromJSON(dataset)
  df <- readlines$objects$person
  
  vals <- substr(df$birthday, 1, 4)
  valsnum <- as.numeric(vals)
  thisyear <- format(Sys.Date(), "%Y")
  thisyearnum <- as.numeric(thisyear)
  age <- thisyearnum - valsnum
  
  df$age <- age
  
  return(df)
}
```

F. Use (call, invoke) the function, and store the results in df


```{r}
df <- addnew()
```

## Part 2: Investigate the resulting dataframe 'df'

A.	How many senators are women? 

sum(df$gender == "female")
#There are 24 female senators

```{r}

```

B.	How many senators have a YouTube account?  

#73 senators have a youtube account

```{r}
sum(!is.na(df$youtubeid))
```

C.	 How many women senators have a YouTube account?
#There is 16 female senators with a youtube

```{r}
womenyoutube <- subset(df, gender=='female' & !is.na(youtubeid))
```

D.	Create a new dataframe called **youtubeWomen** that only includes women senators who have a YouTube account.   


```{r}
womenyoutube <- subset(df, gender=='female' & !is.na(youtubeid))
```

E.	Make a histogram of the **age** of senators in **youtubeWomen**, and then another for the senetors in **df**. Add a comment describing the shape of the distributions.

#Both distributions are just a bit skewed to the left but they both seem to be approavhing normal as the number of senators included goes up

```{r}
hist(womenyoutube$age)
hist(df$age)
```
